CannonRV: Measuring radial velocities with a data-driven spectral model

### 2016-11-26 paper/ms.tex
grammar check

%Can you interpret a, b, c in terms of radial velocity? If so, how?
After obtaining a, b and c for these stars, we can calibrate the errors in the radial velocities of the combined spectra for these stars. These combined spectra are stored in Apstar files, which are generated by different individual visits after resampling them onto a common, logarithmically-spaced wavelength scale, and after removing derived radial velocities of each visit. The apStar files also include the individual visit spectra, resampled and shifted to rest wavelength. If we need to use the individual visits before resampling, apVisit files are needed. In our experiment, we use the resampled spectrum in Apstar file. The radial velocities of the spectrum in Apstar files should be 0, but in fact they are not. This is because of the errors in radial velocity measurement by the APOGEE team. The APOGEE team determine the radial velocities of the stars by two steps.
\bigskip
\\
1. Estimate the radial velocity of each visit by cross-correlating the visit spectrum against a grid of synthetic spectra.
\\
2. Compare the spectrum of each visit with the combined spectrum and re-derive the radial velocity for each visit iteratively. 
\bigskip

After obtaining the velocities, the APOGEE team use the de-Doppler wavelengths and resample the spectrum onto the final logarithmically-spaced wavelength scale using sinc interpolation. Finally, generate combined spectrum by using spectrum of different visits. Both combined spectrum and each visit may have radial velocity errors because the synthetic spectra are not perfect.

@@

Consider Relativistic Doppler effect. the redshift factor is defined as z. The velocity is positive if the star is leaving us. The radial velocities of the stars are much smaller than c, so we have:

@@

The radial velocity is related to parameter a, b and c. First, we need to assume that the inferred spectrum generated by the Cannon is set at rest, which means the radial velocity of it is zero. Second, we think the radial velocity of the data spectrum is equal to the radial velocity of the optimized spectrum. This is because the optimized spectrum is obtained by fitting the inferred spectrum to the data spectrum. 
\bigskip

To begin with, we set the radial velocity to be positive if the star is leaving us. If a=1, b=0 and c=0, the optimized spectrum moves one pixel left from the inferred spectrum, which means the data spectrum should move one pixel right and the radial velocity of the data spectrum is not zero and it has a blue shift. Thus the radial velocity is negative and can be obtained by using (8). It's more convenient to use the logarithmic wavelength.
\begin{equation}
\log_{10}\lambda_0 - \log_{10}\lambda_s = \log_{10}(1+\beta)
\end{equation}

All spectrum in Apstar files are set to the same wavelength scale. The wavelength start at 15100.802 $\AA$ and the logarithmic wavelength grid spacing between two pixels is $6*10^{-6}$. By using (9), we find that one pixel is related to radial velocity 4144.68 m/s, which means if a=1 b= 0 and c=0, the radial velocity error in the APOGEE data is $-4144.68 m/s$

@@

Before deriving a, b and c for each visit, we need to continuum-normalize the spectrum of each visit. This is because the absolute value of each visit is different. Therefore, each visit spectrum needs to be roughly continuum normalized before they are combined. The APOGEE team use a median filter to continuum-normalize the spectrum of each visit. 
\bigskip

The method we use is the same as the one in Ness et al 2015. The spectrum for each visit we use are in Apstar files, which are already resampled to the sample wavelength scale with 8575 pixels.  First, cut all 8575 pixels into three chunks: [[371,3192], [3697,5997], [6461,8255]] to avoid gaps in the spectrum. Then, do pseudo-normalization in these three regions and obtain the pseudo-normalized flux and inverse variance. Third, put a mask on the spectrum. Finally, fit the continuum spectrum in the continuous regions by using a 3-order sinusoid function and obtain the continuum-normalized spectrum. Our method is different from the one used by the APOGEE team and we will demonstrate it works.

To check whether it works, we can infer labels by using the normalized flux of each visit. If the inferred labels from each visit are similar, the normalization works. We define $\triangle Teff$, $\triangle$ logg and $\triangle$ Fe/H as:

@@

Also, we can look at the normalized spectrum of each visit. If the absolute value of normalized flux for each visit is similar to each other, the normalization works.

@@

Here we choose the four stars with biggest a-c and obtain the radial velocity corrections for them. The spectrum we use are continuum-normalized combined spectrum for these four stars. Then we derive their radial velocity corrections and de-Doppler the wavelengths. Use Sinc interpolation to resample them onto the same wavelength scale and obtain the calibrated spectrum. The wavelength scale is the same as the one in Apstar file, which is described in the former section. Finally, plot the data spectrum and the calibrated spectrum in figure 11.

@@

After that, we want to calibrate the spectrum for each visit. To begin with, we need to check whether our normalization works in each visit. The method is to infer labels for each visit and compare them with the ones from the combined spectrum. If our normalization works, the inferred labels from each visit for the same star should be similar to the ones from the combined spectrum. We plot the histogram which shows the deviations of the inferred labels of each visits from the ones generated by the combined spectrum for the same star.  The data we use is the 614 stars in experiment 1 to 4. If most of the deviations are small, the normalized spectrum for each visit doesn't lose shape, which means our normalization method works.

>>>>

%Can you interpret a, b, c in terms of radial velocity? If so, how?
After obtaining a, b and c for these stars, we can calibrate the errors in the radial velocities of the combined spectra of these stars. These combined spectra are stored in Apstar files, which are generated by different individual visits after resampling them onto a common, logarithmically-spaced wavelength scale, and after removing derived radial velocities of each visit. The apStar files also include the individual visit spectra, resampled and shifted to rest wavelength. If we need to use the individual visits before resampling, apVisit files are needed. In our experiment, we use the resampled spectrum in Apstar files. The radial velocities of the spectrum in Apstar files should be 0, but in fact they are not. This is because of the errors in radial velocity measurement by the APOGEE team. The APOGEE team determines the radial velocities of the stars by two steps.
\bigskip
\\
1. Estimate the radial velocity of each visit by cross-correlating the visit spectrum against a grid of synthetic spectra.
\\
2. Compare the spectrum of each visit with the combined spectrum and re-derive the radial velocity for each visit iteratively. 
\bigskip

After obtaining the velocities, the APOGEE team uses the de-Doppler wavelengths and resample the spectrum onto the final logarithmically-spaced wavelength scale by using sinc interpolation. Finally, generate the combined spectrum by using the spectrums of different visits. Both combined spectrum and each visit may have radial velocity errors because the synthetic spectra are not perfect.

@@

Consider Relativistic Doppler effect. The redshift factor is defined as z. The radial velocity is positive if the star is leaving us.

@@

The radial velocity is related to parameters a, b and c. First, we need to assume that the inferred spectrum generated by the Cannon is set at rest, which means the radial velocity of it is zero. Second, we think the radial velocity of the data spectrum is equal to the radial velocity of the optimized spectrum. This is because the optimized spectrum is obtained by fitting the inferred spectrum to the data spectrum. 
\bigskip

To begin with, we set the radial velocity to be positive if the star is leaving us. If a=1, b=0 and c=0, the optimized spectrum moves one pixel left from the inferred spectrum, which means the data spectrum should move one pixel right and the radial velocity of the data spectrum is not zero and it has a blue shift. Thus the radial velocity correction is negative and can be obtained by using (8). It's more convenient to use the logarithmic wavelength.
\begin{equation}
\log_{10}\lambda_0 - \log_{10}\lambda_s = \log_{10}(1+\beta)
\end{equation}

All spectrums in Apstar files are set to the same wavelength scale. The wavelength start at 15100.802 $\AA$ and the logarithmic wavelength grid spacing between two pixels is $6*10^{-6}$. By using (9), we find that one pixel is related to radial velocity 4144.68 m/s, which means if a=1 b= 0 and c=0, the radial velocity error in the APOGEE data is $-4144.68 m/s$

@@

Before deriving a, b and c for each visit, we need to continuum-normalize the spectrum of each visit. This is because the absolute value of each visit is different. Therefore, each visit spectrum needs to be roughly continuum normalized before they are combined. The APOGEE team uses a median filter to continuum-normalize the spectrum of each visit. 
\bigskip

Our method is the same as the one in Ness et al 2015. The spectrum for each visit is in Apstar files, which are already resampled onto the sample wavelength scale with 8575 pixels. First, cut all 8575 pixels into three chunks: [[371,3192], [3697,5997], [6461,8255]] to avoid gaps in the spectrum. Then, do pseudo-normalization in these three regions and obtain the pseudo-normalized flux and inverse variance. Third, put a mask on the spectrum. Finally, fit the continuum spectrum in the continuous regions by using a 3-order sinusoid function and obtain the continuum-normalized spectrum. Our method is different from the one used by the APOGEE team and we will demonstrate it works.

To check whether it works, we can look at the normalized spectrum of each visit. If the absolute value of the normalized flux for each visit is similar to each other, the normalization works.

Also, we can infer labels by using the normalized flux of each visit. If the inferred labels from each visit are similar to the ones from the combined spectrum, the normalization works. We define $\triangle Teff$, $\triangle$ logg and $\triangle$ Fe/H as:

@@

@@

Here we choose the four stars with biggest a-c and obtain the radial velocity corrections for them. The spectrum we use are continuum-normalized combined spectrum of these four stars. Then we derive their radial velocity corrections and de-Doppler the wavelengths. Use Sinc interpolation to resample them onto the same wavelength scale and obtain the calibrated spectrum. The wavelength scale is the same as the one in Apstar file, which is described in the former section. Finally, plot the data spectrum and the calibrated spectrum in figure 11.

@@

After that, we want to calibrate the spectrum for each visit. To begin with, we need to check whether our normalization works in each visit. The method is to infer labels for each visit and compare them with the ones from the combined spectrum. If our normalization works, the inferred labels from each visit for the same star should be similar to the ones from the combined spectrum. We plot the histogram which shows the deviations of the inferred labels of each visit from the ones generated by the combined spectrum of the same star. The data we use is the 614 stars in experiment 1 to 4. If most of the deviations are small, the normalized spectrum for each visit doesn't lose shape, which means our normalization method works.
